---
title: "Linear Regression"
output: rmarkdown::html_vignette
  %\VignetteIndexEntry{Vignette lab4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
require("lab4");
require("ggplot2");
require("png");
require("magick");
linreg<-linreg$new(formula=Petal.Length ~ Species,data=iris)
```


# Linear regression
This package executes a simple linear regression calculations. The functions include:
  * coef(): finds the regression coefficients
  * print(): prints a brief summary of the data
  * plot(): plots the fitted values and residuals
  * summary(): provides a full summary statistics of the data
  * resid(): finds the residuals of estimates
  * pred(): find the prediction of the data based on estimated coefficients.
  

```{r echo=FALSE}
linreg <- setRefClass(Class = "linreg",
                      
                      fields  = list(
                        formula ="formula",
                        data = "data.frame",
                        X = "matrix",
                        y = "matrix",
                        beta_hat = "matrix",
                        y_hat = "matrix",
                        u_hat = "matrix",
                        std_u_hat = "vector",
                        df = "numeric",
                        sigma2 = "numeric",
                        variance = "numeric",
                        variance.beta = "matrix",
                        t_value = "numeric",
                        p_value = "numeric",
                        par = "character"
                      ), 
                      methods = list(
                        
                        initialize = function(formula,data){
                          #model.matrix
                          stopifnot(is.data.frame(data))
                          stopifnot(all.vars(formula) %in% colnames(data))
                          # create informative labels for data sets and plots
                          data <<- data
                          formula <<- formula
                          #Setting up the model
                          X <<- model.matrix(formula, data)
                          #Setting up the dependent variable
                          ycols <-all.vars(formula)[1]
                          y <<- as.matrix(data[ycols])
                          par <<- deparse(substitute(data))
                          beta_hat <<- as.matrix(solve(t(X)%*%as.matrix(X))%*%as.matrix(t(X))%*%as.matrix(y))
                          y_hat <<- as.matrix(as.matrix(X)%*%as.matrix(beta_hat))
                          u_hat <<- as.matrix(y-y_hat) #for calculating the summary stat
                          std_u_hat <<- as.vector(-abs(scale(u_hat))) #standardized residuals
                          df <<- as.numeric(nrow(X)-ncol(X)) #for calculating the summary stat
                          sigma2 <<- as.numeric(as.matrix(t(u_hat)%*%as.matrix(u_hat))/df) #for calculating the summary stat
                          variance <<- as.numeric(sqrt(sigma2))
                          variance.beta <<- as.matrix(sigma2 * solve(as.matrix(t(X)%*%as.matrix((X))))) #finding the variance
                          t_value <<- as.numeric(beta_hat/(sqrt(diag(variance.beta))))
                          p_value <<- as.numeric(2*pt(-abs(t_value),df))#getting the p_values
                        },
                        coef = function(){
                          coef<- as.vector(beta_hat)
                          names(coef) <- rownames(beta_hat)
                          return(beta_hat)
                          },
            
                        print = function(){
                          coef <- numeric()
                          cat("Call:\nlinreg(formula = ",
                              Reduce(paste,deparse(formula)),
                              ", data = ",
                              par,
                              ")\n\nCoefficients:\n ",sep="")
                          
                          for (i in 1:length(beta_hat)){
                            coef[i]<-beta_hat[i]
                          }
                          coef.names<-paste(" ",rownames(beta_hat)[1],"",sep="")
                          for (i in (2:length(coef))){
                            coef.names<-paste(coef.names," ",rownames(beta_hat)[i],sep="")
                          }
                            cat(coef.names)
                            cat("\n")
                            cat(coef,sep="")
                          #cat("\n","Call:","\n",paste("linreg(formula = ", format(formula), ", data = ", parse , ")\n\n ", sep = ""))
                          #print.default(format(coef()), print.gap = 2L,quote = FALSE)
                          #cat("\n")
                        },
                        plot = function(){
                          #The required packages for plotting and setting up the 
                          #background image
                          library("ggplot2")
                          library("png")
                          library("magick")
                          plt_theme <- theme(
                            plot.background = element_rect(fill = '#b8f5f9',linetype = 2),    # Background of the entire plot
            
                            panel.background = element_rect(fill = '#54f4ff'),   # Background of plotting area
                            panel.border = element_rect(colour = "#c2c8d6", fill = NA),       # Border around plotting area.
                            # fill argument should be NA
                            panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                                            colour = "white"),   # Major grid lines
                            panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                                            colour = "white"),   # Minor grid lines
                            axis.line = element_line(colour = "black")
                          )
                          #liu_logo <- image_read_svg("https://upload.wikimedia.org/wikipedia/commons/5/57/Linkoping_University_Logo.svg")
                          dat1 <- data.frame(y_hat, u_hat)
                          pl1<-ggplot(dat1, aes(y_hat, u_hat))+
                            geom_point()+
                            geom_smooth(method="lm", na.rm=TRUE, color="blue")+
                            xlab(paste("Fitted Values"))+
                            ylab("Residuals")+
                            ggtitle("Residuals vs Fitted Plot")+
                            plt_theme
                    
                          dat2 <- data.frame(y_hat, std_u_hat)
                          pl2<-ggplot(dat2, aes(y_hat, std_u_hat))+
                            geom_point()+
                            geom_smooth(method="lm", na.rm=TRUE, color="blue")+
                            xlab(paste("Fitted Values"))+
                            ylab("Residuals")+
                            ggtitle("Standardized Residuals vs Fitted Plot")+
                            plt_theme
                         
                          return(list(pl1,pl2))
                        },
                        resid = function(){
                          return(u_hat)
                        },
                        pred = function(){
                          return(y_hat)
                        },
                        
                        summary = function(){
                          cat("linreg(formula = ", format(formula), ", data = ", par, ") :\n\n ", sep = "")
                          #result<- setNames(object = data.frame(cbind(beta_hat,sqrt(diag(variance.beta)),t_value,round(p_value,2),significance(p_value))),
                            #                nm =c("Estimates", "Std. Error","t value", "p value","Significance"))
                          result<-data.frame(cbind(round(beta_hat,2),round(sqrt(diag(variance.beta)),2),round(t_value,2),round(p_value,2),significance(p_value)))
                          colnames(result) <-c("Estimates", "Std. Error","t value", "p value","Significance")
                          tabl(result)
                          cat("\n Residual standard error: ", variance, " on ", df, " degrees of freedom", sep = "")
                        }
                      )
)
significance <- function(p_value) {
  significance <- ifelse(p_value < 0.001, "***",
                         ifelse(p_value < 0.01, "**",
                                (ifelse(p_value < 0.05, "*",
                                        (ifelse(p_value < 0.1, "."," "))))))
  return(significance)
}
tabl<-function(result){
  print(result)
}

```

The functions that can be implemented by this class are as follows:
** coef()
** print()
** plot()
** resid()
** pred()
** summary()


# Functions
## coef()
```{r echo=FALSE}
coef = function(){
  coef<- as.vector(beta_hat)
  names(coef) <- rownames(beta_hat)
  return(beta_hat)
  }

```
This function calculates and prints out the estimates of coefficients

## print()
```{r echo=FALSE}
print = function(){
                          coef <- numeric()
                          cat("Call:\nlinreg(formula = ",
                              Reduce(paste,deparse(formula)),
                              ", data = ",
                              par,
                              ")\n\nCoefficients:\n ",sep="")
                          
                          for (i in 1:length(beta_hat)){
                            coef[i]<-beta_hat[i]
                          }
                          coef.names<-paste(" ",rownames(beta_hat)[1],"",sep="")
                          for (i in (2:length(coef))){
                            coef.names<-paste(coef.names," ",rownames(beta_hat)[i],sep="")
                          }
                            cat(coef.names)
                            cat("\n")
                            cat(coef,sep="")
                          #cat("\n","Call:","\n",paste("linreg(formula = ", format(formula), ", data = ", parse , ")\n\n ", sep = ""))
                          #print.default(format(coef()), print.gap = 2L,quote = FALSE)
                          #cat("\n")
                        }
```

The print() function prints out the formulation of the regression equation as well as a brief summary of the data and estimations.

## plot()
```{r echo = FALSE}
plot = function(){
                          #The required packages for plotting and setting up the 
                          #background image
                          rlab4::require("ggplot2")
                          rlab4::require("png")
                          rlab4::require("magick")
                          plt_theme <- theme(
                            plot.background = element_rect(fill = '#b8f5f9',linetype = 2),    # Background of the entire plot
            
                            panel.background = element_rect(fill = '#54f4ff'),   # Background of plotting area
                            panel.border = element_rect(colour = "#c2c8d6", fill = NA),       # Border around plotting area.
                            # fill argument should be NA
                            panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                                            colour = "white"),   # Major grid lines
                            panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                                            colour = "white"),   # Minor grid lines
                            axis.line = element_line(colour = "black")
                          )
                          liu_logo <- image_read_svg("https://upload.wikimedia.org/wikipedia/commons/5/57/Linkoping_University_Logo.svg")
                          dat1 <- data.frame(y_hat, u_hat)
                          pl1<-ggplot(dat1, aes(y_hat, u_hat))+
                            geom_point()+
                            geom_smooth(method="lm", na.rm=TRUE, color="blue")+
                            xlab(paste("Fitted Values"))+
                            ylab("Residuals")+
                            ggtitle("Residuals vs Fitted Plot")+
                            plt_theme
                    
                          dat2 <- data.frame(y_hat, std_u_hat)
                          pl2<-ggplot(dat2, aes(y_hat, std_u_hat))+
                            geom_point()+
                            geom_smooth(method="lm", na.rm=TRUE, color="blue")+
                            xlab(paste("Fitted Values"))+
                            ylab("Residuals")+
                            ggtitle("Standardized Residuals vs Fitted Plot")+
                            plt_theme
                         
                          return(list(pl1,pl2))
                        }
```

The plot function creates two plots. The plot of residuls vs. fitted values as well as the standardized residuals vs fitted values.

## resid()
```{r echo=FALSE}
resid = function(){
                          return(u_hat)
}
```

The resid() function prints out the residuals of estimates.

## pred()
```{r echo=FALSE}
pred = function(){ return(y_hat)
  }
```

The pred() function provides the predictions for the variable under analysis.

## summary()
```{r echo=FALSE}
summary = function(){
                          cat("linreg(formula = ", format(formula), ", data = ", par, ") :\n\n ", sep = "")
                          #result<- setNames(object = data.frame(cbind(beta_hat,sqrt(diag(variance.beta)),t_value,round(p_value,2),significance(p_value))),
                            #                nm =c("Estimates", "Std. Error","t value", "p value","Significance"))
                          result<-data.frame(cbind(round(beta_hat,2),round(sqrt(diag(variance.beta)),2),round(t_value,2),round(p_value,2),significance(p_value)))
                          colnames(result) <-c("Estimates", "Std. Error","t value", "p value","Significance")
                          tabl(result)
                          cat("\n Residual standard error: ", variance, " on ", df, " degrees of freedom", sep = "")
                        }
significance <- function(p_value) {
  significance <- ifelse(p_value < 0.001, "***",
                         ifelse(p_value < 0.01, "**",
                                (ifelse(p_value < 0.05, "*",
                                        (ifelse(p_value < 0.1, "."," "))))))
  return(significance)
}
tabl<-function(result){
  print(result)
}
```


The summary() function provides the summary statistics of the data, the p-values and their significance level.

# Tests
We used the data from Iris to create a test suit for our class.

In the following the format that the inputs should be handled is shown.
```{r}
func<-linreg$new(Petal.Length~Sepal.Width+Sepal.Length, data = iris)
func$beta_hat
func$coef()
```

